Galaxy Zoo: Morphological Classifications for 120,000 Galaxies in HST Legacy Imaging

### 2016-05-31 writeup/gz_hubble_data.tex
Switching order of 7.2 and 7.3

@@

All of the Legacy surveys included in the GZH imaging have had morphological
catalogues previously published; these have significant differences in
the number of galaxies, size and magnitude limits, classification scheme, and
the methods used for measuring morphology. These catalogues have been
cross-matched to GZH to compare the results. This is not presented as an
endorsement of any particular method, but as an exploration of the strengths
and weaknesses of the Galaxy~Zoo crowdsourced catalogues compared to products
made with machine learning, automatic fits, and expert visual classification.

@@

Morphologies for GEMS galaxies were measured by \citet{hau07}, who used
single-component \sersic{} fits to the F850LP imaging. A 1\arcsec~positional
match for the GEMS galaxies in Table~9 in \citet{hau07} to the GZH sample gives
8,846~galaxies in both \citet{hau07} and GZH. The primary morphological
parameter in the automated catalogue is the \sersic{} index $n$ defining the
radial surface brightness profile.  ``Elliptical'' galaxies are selected by
$n>2.5$ and ``spirals''  by $n<2.5$.  There is no automatic measurement of
irregular or clumpy structure.

@@

COSMOS galaxies have multiple published datasets automatically classifying
morphology, all using a variation of non-parametric measurements. \citet{cas07}
used a combination of concentration ($C$), asymmetry ($A$), $G$, and $M_{20}$
\citep{cas05} make empirical divisions based on several hundred training images
to assign galaxy morphology. A similar method was employed by \citet{tas11},
using the same non-parametric indices but with a different method of
calculating the Petrosian radius and total light profile, and using a
nearest-neighbors method to label morphologies. \citet[][ZEST]{sca07} used $C$,
$A$, $G$, $M_{20}$, the galaxy ellipticity ($\epsilon$), and \sersic{} index
($n$) to quantify morphology; a principal component analysis is used to assign
galaxies to categories. All three methods use discrete morphological categories
of ellipticals, spirals, and peculiar/irregulars. We use the same categories
with the exception of assigning S0/bulge-dominated disks (type~2.0) from ZEST
into the early-type category.

@@

The AEGIS sample had the fewest number of matched galaxies (485) with GZH;
while the statistical uncertainties are as a result higher, the strongest
difference between spirals and early-types is seen at \fbest$>0.8$. GOODS
galaxies, which are the only comparison sample that also uses visual
morphology, shows a very similar trend, with nearly 100\% pure samples at
the lowest and highest values of \fbest.  Galaxies from GEMS (which are
only split into two types) also have a strong correlation with \fbest,
which is somewhat surprising (but encouraging) considering the known limits
of assigning galaxy morphology based only on a single \sersic{} parameter.

Visual inspection of images of the exceptions --- where both \fbest{} and $n$
are high --- show that most are obvious spirals but with prominent bulges. This
means that the single-component \sersic{} fit is likely choosing too small of
an effective radius and missing the extended disk structure for a large
population of galaxies. This is a known issue with using $n$ as a standalone
measure of galaxy morphology -- it is relatively common, for example, to
measure high $n$ for an object with an exponential disk but a central cusp in
the light profile (eg, due to an AGN, cluster, foreground star, noise, etc).

@@

We repeated all analyses comparing morphologies between GZH and external
pulblications on the entire GZH dataset, rather than the volume-limited sample
($z<1.0,m<22.5$) discussed above. If galaxies at all redshifts and at the
original magnitude limit of $m<23.5$ are used, the correlations between all of
the automatic morphological populations and GZH decrease significantly.  The
exceptions are the GEMS galaxies, which are the only sample that was visually
classified and had a shallower magnitude cut on their catalogue.
Based on these comparisons, we suggest that use of the GZH catalogue be limited
to galaxies with $m_\mathrm{I|i|z}<22.5$; fainter targets at the sensitivity of the \hst{}
Legacy surveys have been demonstrated to be less reliable, regardless of the
method of morphological classification being used.

Finally, there are 7,681~galaxies in the GOODS-S field with morphological
classifications in both GZH and the Galaxy~Zoo:~CANDELS project (Simmons
et~al., submitted). Since both the sensitivity and filters for the two sets of
images are significantly different (and there is no correction for redshift
bias applied to GZC), there is no prior reason to expect a perfect correlation
between the separate vote fractions for the projects. Briefly, we note that the
\ffeatures{} value for GZH is on average higher than GZC; the effect is
strongest at $f_\mathrm{features,GZC}<0.3$, for which roughly half the galaxies
have $f_\mathrm{features,GZH}>0.5$. However, the correlation between vote
fractions is single-valued (although not linear, with a Pearson correlation
coefficient of $r=0.6$), and should be possible to calibrate using a similar
approach to that described in Section~\ref{sec:debiasing}; the correlation
between other tasks, such as edge-on galaxies is significantly stronger
($r=0.9$). So while the raw vote fractions are not directly comparable, the
initial analysis indicates that the broad morphologies are at least consistent.

>>>>

There are 7,681~galaxies in the GOODS-S field with morphological
classifications in both GZH and the Galaxy~Zoo:~CANDELS project (Simmons
et~al., submitted). Since both the sensitivity and filters for the two sets of
images are significantly different (and there is no correction for redshift
bias applied to GZC), there is no prior reason to expect a perfect correlation
between the separate vote fractions for the projects. Briefly, we note that the
\ffeatures{} value for GZH is on average higher than GZC; the effect is
strongest at $f_\mathrm{features,GZC}<0.3$, for which roughly half the galaxies
have $f_\mathrm{features,GZH}>0.5$. However, the correlation between vote
fractions is single-valued (although not linear, with a Pearson correlation
coefficient of $r=0.6$), and should be possible to calibrate using a similar
approach to that described in Section~\ref{sec:debiasing}; the correlation
between other tasks, such as edge-on galaxies is significantly stronger
($r=0.9$). So while the raw vote fractions are not directly comparable, the
initial analysis indicates that the broad morphologies are at least consistent.

@@

All of the Legacy surveys included in the GZH imaging have had morphological
catalogues previously published; these have significant differences in
the number of galaxies, size and magnitude limits, classification scheme, and
the methods used for measuring morphology. These catalogues have been
cross-matched to GZH to compare the results. 
%This is not presented as an
%endorsement of any particular method, but as an exploration of the strengths
%and weaknesses of the Galaxy~Zoo crowdsourced catalogues compared to products
%made with machine learning, automatic fits, and expert visual classification.

@@

Morphologies for GEMS galaxies were measured by \citet{hau07}, who used
single-component \sersic{} fits to the F850LP imaging. There are 8,846~galaxies
with measurements in both \citet{hau07} and GZH. The primary morphological
parameter in the automated catalogue is the \sersic{} index $n$ defining the
radial surface brightness profile.  ``Elliptical'' galaxies are selected by
$n>2.5$ and ``spirals''  by $n<2.5$.  There is no automatic measurement of
irregular or clumpy structure.

@@

COSMOS galaxies have multiple published datasets automatically classifying
morphology, all using a variation of non-parametric measurements. \citet{cas07}
used a combination of concentration ($C$), asymmetry ($A$), $G$, and $M_{20}$
\citep{cas05} make empirical divisions based on several hundred training images
to assign galaxy morphology. A similar method was employed by \citet{tas11},
using the same non-parametric indices but with a different method of
calculating the Petrosian radius and total light profile, and using a
nearest-neighbors method to label morphologies. \citet[][ZEST]{sca07} used $C$,
$A$, $G$, $M_{20}$, the galaxy ellipticity ($\epsilon$), and \sersic{} index
($n$) to quantify morphology; a principal component analysis is used to assign
galaxies to categories. All three methods use discrete morphological categories
of ellipticals, spirals, and peculiar/irregulars. We use the same categories
with the exception of assigning S0/bulge-dominated disks (type~2.0) from ZEST
into the unfeatured category.

@@

The AEGIS sample had only 485~galaxies matched with GZH, and so the statistical
uncertainties in the analysis are higher. However, at \fbest$>0.8$ there is a
clear separation of featured and unfeatured galaxies, though this is not seen
at lower \fbest.  GOODS galaxies, which are the only comparison sample that
also uses visual morphology, shows a very similar trend, with nearly 100\% pure
samples at the lowest and highest values of \fbest.

Galaxies from GEMS (which
are only split into two types) also have a strong correlation with \fbest,
which is somewhat surprising (but encouraging) considering the known limits of
assigning galaxy morphology based only on a single \sersic{} parameter. 
Visual inspection of images of the exceptions --- where both \fbest{} and $n$
are high --- show that most are obvious spirals but with prominent bulges. This
means that the single-component \sersic{} fit is likely choosing too small of
an effective radius and missing the extended disk structure for a large
population of galaxies. This is a known issue with using $n$ as a standalone
measure of galaxy morphology -- it is relatively common, for example, to
measure high $n$ for an object with an exponential disk but a central cusp in
the light profile (eg, due to an AGN, cluster, foreground star, noise, etc).

@@

We repeated all analyses comparing morphologies between GZH and external
pulblications on the entire GZH dataset, rather than the volume-limited sample
($z<1.0,m<22.5$) discussed above. If galaxies at all redshifts and at the
original magnitude limit of $m<23.5$ are used, the correlations between all of
the automatic morphological populations and GZH decrease significantly.  The
exceptions are the GEMS galaxies, which are the only sample that was visually
classified and had a shallower magnitude cut on their catalogue.  Based on
these comparisons, we suggest that use of morphological catalogues with these
\hst{} images be limited to galaxies with $m_\mathrm{I|i|z}<22.5$; fainter
targets have been demonstrated to be less reliable, regardless of the method of
morphological classification being used.
